Operational Loss with Correlated Frequency and Severity: an Analytical Approach

### 2017-07-25 ResponseToReferee.tex

\documentclass{article}
\usepackage{amsmath}
\usepackage{fancyhdr}
\usepackage{amsthm}
\usepackage{caption}
\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage{cite}
\usepackage{float}
\theoremstyle{definition}
\newtheorem{theorem}{Theorem}

\linespread{1.3}
\newcommand{\adjustHeight}{\setlength{\abovecaptionskip}{-15pt}} 
\setlength{\belowcaptionskip}{-4pt}

\makeatletter
\renewcommand\@biblabel[1]{}
\setlength{\parindent}{0pt}

\begin{document}

\title{Response to Referee}
\date{}
\maketitle
I thank the Referee for his or her constructive comments.  I have made significant changes in response to points 1, 3, and 4 from the ``concerns'' section of the Referee Report.  These changes will be enumerated in detail below.

\begin{enumerate}
\item ``One of my concerns is with the motivation. The opening paragraph in the Introduction section makes a convincing argument that the occurrence of operational risk events may be correlated in time due to declining internal controls caused by a significant loss event. However, the second paragraph also talks about frequency being correlated with severity, but the intuition (or empirical evidence) is not presented to motivate such conjecture.''
\\
\\
My original example in the introduction captures both the auto-correlation and the correlation between severity and frequency.  However, this was not clear from the text.  I have revised my introduction to include the following verbiage: ``Following a particularly severe loss event, the manpower required to deal with the event may cause a failure in internal control leading to an in-creased likelihood of an additional event.  The occurrence of an event may be the result of declining internal controls, which may indicate an increased likelihood of  additional  events.   In  such  a  scenario  the  severity  of  an  event  impacts  the frequency of future events, driving correlation between frequency and severity distributions.''
\item ``The paper is quite technical, and for an average reader it is easy to get lost in the technicalities. I have not checked the math, but it looks solid.''
\\
\\
See my response to item \ref{item5}.

\item 
``For the severity distribution, the paper uses alpha-stable distributions that are known to exhibit very heavy tails (in addition to their lack of closed-form expression of the density function). I wonder whether the authors can apply their model with more “traditional” severity distributions commonly used in operational risk literature, such as Log-normal, Weibull, and others. Right now, the application of the proposed approach is too narrowly focused, and the analysis would be much richer if a wider variety of severity distributions were used and the results were compared.''
\\
\\
I agree with the Referee.  I have added analysis around the Gamma and Inverse Gaussian distributions as well.  Note that Log-normal and Weibull distributions do not have closed form characteristic functions and so cannot be efficiently used in this modeling framework.  The analysis of these additional distributions has resulted in an expanded ``Severity Distribution'' section and an expanded ``Numerical Implementation'' section including impacts on value at risk for each severity distribution.

\item
``The proposed model leads to increased tail risk that arises from the dependencies between the frequency and severity distributions. Because of this, as shown in Table 1, VaR estimates increase dramatically with the correlation parameter. With this in mind, what incentive would banks have to use the proposed approach if it leads to VaR estimates that are much higher than in the case with zero correlation? Back to my previous point, perhaps VaR would be less sensitive to correlation if other (less heavy-tailed) severity distributions were used.''
\\
\\
I have added verbiage in my conclusions around why firms would want to use this model: ``The
flexibility  of  the  model  allows  firms  to  more  precisely  estimate  the  capital  to hold  for  operational  risk  and  design  superior  controls  to  mitigate  operational risk.   Since  the  LDA  model  is  a  nested  model,  firms  can  also  test  whether the  LDA  modeling  approach  is  robust  to  correlations  between  frequency  and severity and auto-correlations between frequency.''  Additionally, I have tested the impact of the correlation on the other severity distributions.

\item \label{item5}
``The paper as it stands is theoretical. The analysis would benefit significantly if the proposed methods could be tested with real operational loss data (or at least some simulated data) and VaR estimates under the traditional LDA model (zero correlation) and the proposed model were compared and the implications were discussed.''
\\
\\
My paper is intended to be theoretical.  It provides flexible and efficient extensions to the LDA model.  The model will perform at least as well (in sample) as the standard LDA model regardless of the data since the LDA is a nested model.  The output of the model is a distribution.  If I simulated loss data from the model in this paper I would be essentially guaranteed a perfect fit.

\end{enumerate}

\end{document}

### 2021-08-24 main.tex
updated to clarify variables

\section{Model Description}
\subsection{Constant Frequency}
Consider a business with operational loss events.  The random variable that counts the number of loss events in a time period \([0, t]\) is \(N_t\) and follows a Poisson distribution with parameter \(\lambda t\).  The total loss from these loss events is \(X_t=\sum_{j=1} ^ {N_t} L_j\) where each \(L_j\) is independently and identically distributed positive random variables with finite expectation.  These \(L_j\) represent the dollar loss from loss event \(j\).  Since each \(L_j\) is stochastic, \(X_t\) is a compound Poisson jump process.  Intuitively, loss events arrive as a ``stream'' instead of at a point in time, with no two events occurring simultaneously.  For notational convenience, \(L_j\) will denote random draws from the relevant distribution while \(L\) will denote the random variable which the relevant distribution describes.  
\\
\\
There is no general analytic formula for the density of a compound Poisson jump process; however the characteristic function has a simple solution.  Recall that the characteristic function of a random variable \(Z\) is defined as \(\mathbb{E}[e^{\mathrm{i}Z}]\) where \(i\) is the number which satisfies \(i^2=-1\).  Hence a compound Poisson process's characteristic function is derived as follows: 
\[\mathbb{E}[e^{u\mathrm{i}X_t}]=\mathbb{E}\left[\mathbb{E}[e^{u\mathrm{i}X_t}|\mathcal{F}_t]\right]=\mathbb{E}\left[\mathbb{E}\left[\prod_{j=1} ^ {N_t} e^{u\mathrm{i} L_j}|\mathcal{F}_t\right]\right]\]
\[=\mathbb{E}\left[\prod_{j=1} ^ {N_t} \mathbb{E}[e^{u\mathrm{i}L}] \right]=\mathbb{E}\left[e^{\mathrm{ln}\left(\mathbb{E}[e^{u\mathrm{i}L}]\right) N_t}\right]\]
\[=e^{\lambda t \left(\mathbb{E}\left[ e^{\mathrm{i}uL}\right]-1\right)}=e^{t \int_{\mathbb{R}} (e^{\mathrm{i}uL}-1) \lambda d\mu}\]

Where the second line is justified by the moment generating function of a Poisson random variable, \(\mathcal{F}_t\) is the filtration generated by \(N_t\), and \(d\mu\) is the density function of \(L\).  From the characteristic function, it is clear that the jump diffusion is defined by the product \(\lambda d\mu\) where \(\lambda\) controls the frequency of jumps and \(d\mu\) controls the jump size. Making this dependence explicit, the characteristic function of a compound Poisson process is denoted
\begin{equation}\phi_P (u;\,\lambda d\mu, t)=e^{t \int_{\mathbb{R}} (e^{\mathrm{i}uL}-1) \lambda d\mu}\end{equation}
At this point, there is no assumption of correlation between frequency and severity: indeed, the frequency is determined by the constant \(\lambda\).  
\\
\\

@@

 Letting \(\bar{k}=\frac{\delta \lambda \mathbb{E}[L]}{a}+1\),
\begin{equation}\label{tmc}v_t=v_0+\int_0 ^t a(1-\bar{k}v_s)ds+\sigma \int_0 ^ t \sqrt{v_s} dW_s +\delta \sum_{j=1} ^ {N_\tau} L_j \end{equation}

@@

Modeling \(v_t\) in such a manner has several advantages.  First, by Duffie et al. (2000), \(v_t\) has a semi-analytic moment generating function.  Second, \(v_t\) is positive.  This is necessary as time cannot run backward.  Finally, \(v_t\) is mean reverting.  In the long run \(v_t\) will have an expected value with converges to a one; implying that stochastic time is a long run unbiased estimator of calendar time.  The parameter \(a>0\) controls the speed of reversion to the mean: a higher value indicates a quicker return to ``normal'' levels following an innovation.  The parameter \(\sigma>0\) controls the volatility of the diffusion.  The parameter \(\delta>=0\) controls the impact that a jump has on the diffusion.  A \(\delta\) of \(0\) means that there is no impact (and frequency and severity are uncorrelated).  The ``standard'' LDA model (defined as \(\phi_X(u)=e^{t\lambda \left(\mathbb{E}\left[e^{\mathrm{i}uL}\right]-1\right)}\)) is recovered when \(a=\delta=\sigma=0\).

@@

\begin{theorem}
Let \(Y_t\) be a compound Poisson process characterized by \(\lambda d\mu\).  Define the martingale

@@

Then under the probability measure \(\hat{d\mathbb{P}}=\eta_t d\mathbb{P}\), \(Y_t\) is a compound Poisson process characterized by \(e^{u\mathrm{i}L}\lambda d\mu\).

@@

\[=e^{t \int_{R} \left(e^{\mathrm{i}zL}-1\right)\lambda e^{\mathrm{i}uL}d\mu}=\phi_{P} (z;\,e^{u\mathrm{i}L}\lambda d\mu, t)\]

@@

Carr and Wu (2004) prove that this theorem remains applicable with \(\tau\) substituted for \(t\).  This theorem implies that under this new measure, the dynamics of \(v\) are as follows:
\begin{equation}\label{tmcLN}v_t=v_0+\int_0 ^t a(1-\bar{k}v_s)ds+\sigma \int_0 ^ t \sqrt{v_s} dW_s +\delta \sum_{j=1} ^ {\hat{N}_\tau} \hat{L}_j \end{equation}

By Duffie et al. (2000), for the affine process \(v_t\), the following statement holds:
\[\mathbb{E} \left[e^{-r\int_0 ^ {T} v_s ds}\right]=e^{\gamma(0, T)+\zeta(0, T)v_0}\]
Where \(\gamma,\,\zeta\) satisfy

\[
\begin{cases}
\frac{\partial \zeta}{\partial t}=a(\bar{k}+1) \zeta+r-\frac{1}{2} \sigma^2 \zeta^2-\int_{\mathbb{R}} \left(e^{\delta \zeta L}-1\right)\lambda e^{u\mathrm{i}L} d\mu & \zeta(T, T)=0\\
\frac{\partial \gamma}{\partial t}=-a \zeta & \gamma(T, T)=0
\end{cases} \]

In this case, \(r=\lambda(1-\mathbb{E}[e^{u\mathrm{i}L}])\).  The system of ODEs thus reads

\begin{equation}\label{ode}
\begin{cases}
\frac{\partial \zeta}{\partial t}=a(\bar{k}+1) \zeta+\lambda-\lambda\phi_L(u-\mathrm{i}\delta \zeta )-\frac{1}{2} \sigma^2 \zeta^2 & \zeta(T, T)=0\\
\frac{\partial \gamma}{\partial t}=-a \zeta & \gamma(T, T)=0
\end{cases} \end{equation}
Where \(\phi_L(z)\) is the characteristic function of \(L\).

>>>>

\section{Model Description}
\subsection{Constant Frequency}
Consider a business with operational loss events.  The random variable that counts the number of loss events in a time period \([0, t]\) is \(N_t\) and follows a Poisson distribution with parameter \(\lambda t\).  The total loss from these loss events is \(X_t=\sum_{j=1} ^ {N_t} L_j\) where each \(L_j\) is independently and identically distributed positive random variables with finite expectation.  These \(L_j\) represent the dollar loss from loss event \(j\).  Since each \(L_j\) is stochastic, \(X_t\) is a compound Poisson jump process.  Intuitively, loss events arrive as a ``stream'' instead of at a point in time, with no two events occurring simultaneously.  For notational convenience, \(L_j\) will denote random draws from the relevant distribution while \(L\) will denote the random variable which the relevant distribution describes.  
\\
\\
There is no general analytic formula for the density of a compound Poisson jump process; however the characteristic function has a simple solution.  Recall that the characteristic function of a random variable \(Z\) is defined as \(\mathbb{E}[e^{\mathrm{i}Z}]\) where \(i\) is the number which satisfies \(i^2=-1\).  Hence a compound Poisson process's characteristic function is derived as follows: 
\[\mathbb{E}[e^{u\mathrm{i}X_t}]=\mathbb{E}\left[\mathbb{E}[e^{u\mathrm{i}X_t}|\mathcal{F}_t]\right]=\mathbb{E}\left[\mathbb{E}\left[\prod_{j=1} ^ {N_t} e^{u\mathrm{i} L_j}|\mathcal{F}_t\right]\right]\]
\[=\mathbb{E}\left[\prod_{j=1} ^ {N_t} \mathbb{E}[e^{u\mathrm{i}L}] \right]=\mathbb{E}\left[e^{\mathrm{ln}\left(\mathbb{E}[e^{u\mathrm{i}L}]\right) N_t}\right]\]
\[=e^{\lambda t \left(\mathbb{E}\left[ e^{\mathrm{i}uL}\right]-1\right)}=e^{t \int_{\mathbb{R}} (e^{\mathrm{i}ux}-1) \lambda v(dx)}\]

Where the second line is justified by the moment generating function of a Poisson random variable, \(\mathcal{F}_t\) is the filtration generated by \(N_t\), and \(v\) is the density function of \(L\).  From the characteristic function, it is clear that the jump diffusion is defined by the product \(\lambda v(dx)\) where \(\lambda\) controls the frequency of jumps and \(v(dx)\) controls the jump size. Making this dependence explicit, the characteristic function of a compound Poisson process is denoted
\begin{equation}\phi_P (u;\,\lambda v(dx), t)=e^{t \int_{\mathbb{R}} (e^{\mathrm{i}uL}-1) \lambda v(dx)}\end{equation}
At this point, there is no assumption of correlation between frequency and severity: indeed, the frequency is determined by the constant \(\lambda\).  
\\
\\

@@

 %Letting \(\bar{k}=\frac{\delta \lambda \mathbb{E}[L]}{a}+1\),
 Letting \(\rho=\delta \lambda \mathbb{E}[L]\),
\begin{equation}\label{tmc}v_t=v_0+\int_0 ^t a\left(1-\left(1+\frac{\rho}{a}\right)v_s\right)ds+\sigma \int_0 ^ t \sqrt{v_s} dW_s +\frac{\rho}{\mathbb{E}[L] \lambda} \sum_{j=1} ^ {N_\tau} L_j \end{equation}

@@

Modeling \(v_t\) in such a manner has several advantages.  First, by Duffie et al. (2000), \(v_t\) has a semi-analytic moment generating function.  Second, \(v_t\) is positive.  This is necessary as time cannot run backward.  Finally, \(v_t\) is mean reverting.  In the long run \(v_t\) will have an expected value with converges to a one; implying that stochastic time is a long run unbiased estimator of calendar time.  The parameter \(a>0\) controls the speed of reversion to the mean: a higher value indicates a quicker return to ``normal'' levels following an innovation.  The parameter \(\sigma>0\) controls the volatility of the diffusion.  The parameter \(\rho>=0\) controls the impact that a jump has on the diffusion.  A \(\rho\) of \(0\) means that there is no impact (and frequency and severity are uncorrelated).  The ``standard'' LDA model (defined as \(\phi_X(u)=e^{t\lambda \left(\mathbb{E}\left[e^{\mathrm{i}uL}\right]-1\right)}\)) is recovered when \(a=\rho=\sigma=0\).

@@

\begin{theorem}\label{theorem1}
Let \(Y_t\) be a compound Poisson process characterized by \(\lambda v(dx)\).  Define the martingale

@@

Then under the probability measure \(\hat{d\mathbb{P}}=\eta_t d\mathbb{P}\), \(Y_t\) is a compound Poisson process characterized by \(e^{u\mathrm{i}L}\lambda v(dx)\).

@@

\[=e^{t \int_{R} \left(e^{\mathrm{i}zx}-1\right)\lambda e^{\mathrm{i}uL}v(dx)}=\phi_{P} (z;\,e^{u\mathrm{i}x}\lambda v(dx), t)\]

@@

Carr and Wu (2004) prove that this theorem remains applicable with \(\tau\) substituted for \(t\).  This theorem implies that under this new measure, the dynamics of \(v\) are as follows:
\begin{equation}\label{tmcLN}v_t=v_0+\int_0 ^t a\left(1-\left(\frac{\rho}{a}+1\right)v_s\right)ds+\sigma \int_0 ^ t \sqrt{v_s} dW_s +\frac{\rho}{\mathbb{E}[L] \lambda} \sum_{j=1} ^ {\hat{N}_\tau} L_j \end{equation}
Where \(\hat{N}_\tau\) is the time-changed counting process under the ``leverage-neutral'' measure.   Using By Duffie et al. (2000), for the affine process \(v_t\), the following statement holds:
\[\mathbb{E} \left[e^{-r\int_0 ^ {T} v_s ds}\right]=e^{\gamma(0, T)+\zeta(0, T)v_0}\]
Where \(\gamma,\,\zeta\) satisfy

\[
\begin{cases}
\frac{\partial \zeta}{\partial t}=a\left(1+\frac{\rho}{a}\right) \zeta+r-\frac{1}{2} \sigma^2 \zeta^2-\mathbb{\hat{E}}\left[ e^{\zeta \delta L}-1\right] & \zeta(T, T)=0\\
\frac{\partial \gamma}{\partial t}=-a \zeta & \gamma(T, T)=0
\end{cases} \]

By theorem \ref{theorem1}, \(\mathbb{\hat{E}}\left[ e^{ \frac{\rho \zeta L}{\mathbb{E}[L] \lambda}}-1\right]=\int_{\mathbb{R}} \left(e^{\frac{\rho \zeta x}{\mathbb{E}[L] \lambda}}-1\right)\lambda e^{u\mathrm{i}x} v(dx)\).  Since the time-changed characteristic function is \(\mathbb{\hat{E}}\left[e^{\psi(u)\tau} \right] \),  \(r=-\psi(u)=\lambda(1-\mathbb{E}[e^{u\mathrm{i}L}])\).  Making these subsitutions, the system of ODEs becomes

\begin{equation}\label{ode}
\begin{cases}
\frac{\partial \zeta}{\partial t}=(a+\rho) \zeta-\lambda\left(\phi_L(u-\mathrm{i}\frac{\rho}{\mathbb{E}[L] \lambda} \zeta )-1\right)-\frac{1}{2} \sigma^2 \zeta^2 & \zeta(T, T)=0\\
\frac{\partial \gamma}{\partial t}=-a \zeta & \gamma(T, T)=0
\end{cases} \end{equation}
Where \(\phi_L(z)\) is the characteristic function of \(L\).
